Does a present bias influence exploratory choice?

### 2017-02-01 doc/cogsci2017/paper.tex
making cogsci paper edits suggested by Todd

Notably, though, while over-exploration does occur, a bias towards under-exploration in particular seems to be
observed across a wide range of domains. One of the most severe
expressions of under-exploration is learned helplessness, in which an organism
experiences the absence of contol over the environment, learns that the
environment is uncontrollable, and thus ceases to take actions that might allow
it to discover that it can in fact exert control. Learned helplessness has been proposed to
underly most notably depression \citep{Abramson1978, Abramson1989} but also
problems ranging from difficulties in school \citep{Diener1978} to poverty \citep{Evans2005}.
While the cognitive appraisal of experienced events affects the development of
learned helplessness \citep{Abramson1978}, patterns of exploration clearly play
a role as well \citep{Huys2009, Teodorescu2014a}. In the case of depression,
interventions aimed at increasing the exploration of activities this might
produce a sense of pleasure and mastery have been found to be as effective as
those with a more cognitive orientation \citep{Jacobson1996}.

\hl{This is good but kind of goes deep into these topics which aren't that central
to the current paper... i.e., learned helplessness per-se.  it can be a little distracting
and adds a lot of citations.  incidently the underexploration of product categories
seems relevant as well}

@@

There are many other areas in which under-exploration is less clearly
established, but is suspected to play a role in maladaptive behavior.
Insufficient exploratory interaction with outgroups may be one cause of
stereotypes and prejudice \citep{Denrell2005most}, and interventions that
increase inter-group contact reduce stereotypes \citep{Shook2008}. The crowding
out of exploration by exploitation is a concern in organizational behavior as
well \citep{March1991, Levinthal1993}, prompting research into organizational
structures that may preserve exploration \citep{Fang2010}.

@@

Under-exploration is often proposed to be the result of myopia: the
underweighting of temporally distant rewards relative to close ones \citep{March1991, Levinthal1993}. 
\hl{does this someone undermine the novelty of your framing/prediction?}
While
choosing actions that are expected to immediately maximize reward is inherently
exploitative, exploration often requires the decision-maker to suffer short-term
costs in search of better future options.

@@

The the second form of myopia is that the decision-maker
is aware of the future but simply doesn't care as much about it.
A wealth of studies have shown that this form of myopia is an important factor
in human decision making, and that people are strongly
biased towards  the present. For example, people will often prefer a larger, later monetary
reward to a smaller, sooner reward when both rewards are in the future, but will
switch their preference when the time until both rewards is reduced so that the
sooner reward is immediate \citep{Kirby1995}. With monetary rewards, the delay
or speed-up required to observe preference reversals is usually several days. With non-monetary rewards,
such as the cessation of an annoying noise \citep{Solnick1980}, watching a video
when bored \citep{Navarick1998}, or drinking soda
when thirsty \citep{Brown2009}, a bias towards immediate rewards has been observed on the scale of
minutes or even seconds.

@@

@@

Thirty-one participants completed the experiment, which was conducted over
Amazon Mechanical Turk (AMT) using the psiTurk framework \citep{Gureckis2015a}.
Participants were paid \$5.00 for their participation, with a performance-based
bonus of up to \$3.00. All participants received the full \$3.00 bonus.
\hl{Maybe mention randomization proceedure to the two conditions?}

@@

\hl{Complex task to follow...  it might make sense to overview it and then
divide into subsections.  like "consuption tasks" "choice tasks", etc...}

Participants were informed that their job was to perform a monotonous slider task that
would be split into 30-second ``work periods'', but that they would be able to
make choices throughout the experiment that would give them a chance to watch a
Youtube video instead. The number of remaining work periods in the experiment was shown
at the top of the screen, as was the number of seconds left in the current work period.

The slider task was based of a task previously used by \citet{Gill2012}. In each
period of the slider task, five horizontal sliders appeared on the screen (Figure~\ref{fig:taskpictures}\emph{a}). Each
started at a random setting between 0 and 100, with the slider's value
shown to its right, with a random horizontal offset so that the
sliders were not aligned. The participant's task was to use the mouse to move
each slider to ``50'' before the work period ended. When a participant released
the mouse at the correct setting, the slider turned green to show it had been
completed. To ensure that the task took close to the allotted 30 seconds, at the
beginning of the task only the top slider was enabled, and the other four were
grayed out. Additional sliders were enabled at five-second intervals, such that
all five sliders were available after 20 seconds.

Before beginning the experiment, participants chose one of four videos available
on Youtube: an episode of ``Planet Earth'', and episode of ``The Great British
Bakeoff'', and episode of ``Mythbusters'', or an Ellen Degeneres comedy special.
The video was embedded in the experiment with all user controls (such as
skipping ahead) disabled (Figure~\ref{fig:taskpictures}\emph{b}). When given access to the video, participants had to keep the
browser window open and hold down the space bar for the video to play. This allowed us to ensure that
participants maintained engagement with the content.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/machinemappictures.png}
\caption{The machine display seen by participants to allow them to keep track of
  the value of each of the six machines and the next time it would be ready to
  make a choice or produce an outcome. \emph{a}: the display seen by
  participants in Experiment 1 and the immediate condition of Experiment 2.
  \emph{b}: the display seen by participants in teh delayed condition of
  Experiment 2. \hl{I wonder if you draw and arrow to show the movement of the machines in the queue?}}
\label{fig:machinemappictures}
\end{figure*}

In the main part of the experiment, participants completed a total of 70 work
periods. For the first 10 work periods, participants simply clicked a button to begin the
slider task. After these initial periods, participants gained access to six
``machines'', each of which had a ``current spinner'' and a ``new spinner'' (Figure~\ref{fig:taskpictures}\emph{c}).
Before each work period, participants clicked the circled machine, \hl{Unclear what circled machine means}
were shown the machine and made a choice about which spinner to run. Based on their choices, there was a
chance the machine would perform the work task for the participant, allowing the
participant to watch their chosen video instead.
\hl{I wonder if you just say more directly that you spin either an existing dial (exploit) or a new
dial (explore).  If the machine came up yellow it would do the "work" for you and you could watch
videos.  If not they you had to do the slider tasks yourself}

The machine ran immediately after the choice was
made, and affected the next work period, as shown in
Figure~\ref{fig:machinetimeline}\emph{a} and Figure~\ref{fig:machinemappictures}\emph{a}. It then ``cooled off'' for the
following five periods, as choices were made with the other five machines.

At each choice, participants had to select between two options: run the
machine's ``current spinner'' or run a ``new spinner''. Each machine had two
circular spinners with arrows at the top. The current spinner was split into
five black and five gold wedges. If a participant chose the
current spinner, it spun (immediately or after four periods, depending
on the condition) and, if the arrow landed on the gold wedge, the machine
worked and the participant could watch the video. Initially, the
current spinner's gold wedges were randomly set for each machine to cover between
1/3 and 2/3 of the spinner.

The new spinner initially showed a question mark. If a participant chose the
new spinner, then (immediately or after four periods) a new spinner was
created and appeared on the machine. In the high-explore condition, the new spinner's gold wedges could cover
anywhere from 25\% of the spinner to 100\% of the spinner. In the low-explore condition, the new spinner's gold wedges could cover
anywhere from 0\% of the spinner to 75\% of the spinner. The new spinner then
spun and, if the arrow landed on a gold wedge, the machine worked.
Additionally, if the new spinner had a greater gold area than the current
spinner, the new spinner was ``saved'' and the current spinner was
updated to the new spinner.  \emph{This meant that
while choosing a new spinner carried risk, it could carry long-term
benefits as the current spinner could be improved from its initial value.}
Importantly, the value of choosing a new spinner was higher in the high-spinner
condition, because the range of outcomes was shifted towards higher values.

@@

During the machine choices, participants had access to an ``info'' button at the
bottom of the screen that provided reminders about the dynamics of the task. In
addition, before beginning the full experiment, participants completed two
practice phases. First, they performed several trials of practice using the
machines, with the actual work periods removed. Then, they performed three work
periods practicing the slider task.

@@

\hl{A little discussion here might help frame things}

@@

Forty participants completed the experiment, which was conducted over
AMT using the psiTurk framework.
Participants were paid \$5.00 for their participation, with a performance-based
bonus of up to \$3.00. All participants received the full \$3.00 bonus.
\hl{Randomization to conditions}

@@

Experiment 2 followed the design of Experiment 1, with the following changes. In
both conditions, the possible winning proportion of a new spinner ranged from 0\% to 100\% of
the spinner. In the immediate condition the machine ran immediately after the
choice was made, as in Experiment 1. In the delayed condition, each machine was
presented to the participant four work periods before it was schedules to run,
and the participant made a choice at that time. The machine then had to
``process'' for four work periods, thus delaying
the outcome by over 2 minutes. The participant then clicked the machine again to
return to it and observe its outcome, and either perform the slider task or
watch the video. The machine then cooled of for a single period before being
ready for another choice. The timeline for the delayed condition is shown in
Figure~\ref{fig:machinetimeline}\emph{b}.

@@

\begin{small}
  \noindent
\textbf{Acknowledgments} 
\end{small}

>>>>

Notably, though, while over-exploration does occur, a bias towards under-exploration in particular seems to be
observed across a wide range of domains. In learned helplessness, for example, an organism
experiences the absence of contol over the environment, learns that the
environment is uncontrollable, and thus ceases to take actions that might allow
it to discover that it can in fact exert control. % Learned helplessness has been proposed to
% underly most notably depression \citep{Abramson1978, Abramson1989} but also
% problems ranging from difficulties in school \citep{Diener1978} to poverty \citep{Evans2005}.
While the cognitive appraisal of experienced events affects the development of
learned helplessness \citep{Abramson1978}, patterns of exploration clearly play
a role as well \citep{Huys2009, Teodorescu2014a}. % In the case of depression,
% interventions aimed at increasing the exploration of activities this might
% produce a sense of pleasure and mastery have been found to be as effective as
% those with a more cognitive orientation \citep{Jacobson1996}.

@@

In many other areas under-exploration is less clearly
established, but is suspected to play a role in maladaptive behavior.
Insufficient exploratory interaction with outgroups may be one cause of
stereotypes and prejudice \citep{Denrell2005most}, and interventions that
increase inter-group contact reduce stereotypes \citep{Shook2008}. The crowding
out of exploration by exploitation is a concern in organizational behavior as
well \citep{March1991, Levinthal1993}, prompting research into organizational
structures that may preserve exploration \citep{Fang2010}.

@@

Past theoretical work has proposed that Under-exploration may be the result of myopia: the
underweighting of temporally distant rewards relative to close ones \citep{March1991, Levinthal1993}. 
% \hl{does this someone undermine the novelty of your framing/prediction?}
While
choosing actions that are expected to immediately maximize reward is inherently
exploitative, exploration often requires the decision-maker to suffer short-term
costs in search of better future options.

@@

The second form of myopia is that the decision-maker
is aware of the future but simply doesn't care as much about it.
A wealth of studies have shown that this form of myopia is an important factor
in human decision making, and that people are strongly
biased towards  the present. For example, people will often prefer a larger, later monetary
reward to a smaller, sooner reward when both rewards are in the future, but will
switch their preference when the time until both rewards is reduced so that the
sooner reward is immediate \citep{Kirby1995}. With monetary rewards, the delay
or speed-up required to observe preference reversals is usually several days. With non-monetary rewards,
such as the cessation of an annoying noise \citep{Solnick1980}, watching a video
when bored \citep{Navarick1998}, or drinking soda
when thirsty \citep{Brown2009}, a bias towards immediate rewards has been observed on the scale of
minutes or even seconds.

@@

In our experimental paradigm, participants made a series of explore--exploit
decisions, where each decision determined whether the following 30-second
consumption task would be unpleasant (moving sliders) or pleasant (watching a
Youtube video). In Experiment 1, we sought to confirm that participants prefered
the video task to the slider task and would seek to increase video-watching
time, and that people were sensitive to explicit, exogenous changes to the value
of exploration. The experiment consisted of two conditions---in the high-explore
condition, the average payoff of exploration was high, and in the low-explore
condition, it was low. We predicted that participants in the high-explore condition
would make more exploratory choices.

@@

Thirty-one participants completed the experiment, which was conducted over
Amazon Mechanical Turk (AMT) using the psiTurk framework \citep{Gureckis2015a}.
Participants were paid \$5.00 for their participation, with a performance-based
bonus of up to \$3.00. All participants received the full \$3.00 bonus.
Participants were pseudo-randomly counterbalanced across two conditions.

@@

% \hl{Complex task to follow...  it might make sense to overview it and then
% divide into subsections.  like "consuption tasks" "choice tasks", etc...}

\emph{Consumption tasks.}
Participants were informed that their job was to perform a monotonous slider task that
would be split into 30-second ``work periods'', but that they would be able to
make choices throughout the experiment that would give them a chance to watch a
Youtube video instead. The number of remaining work periods in the experiment was shown
at the top of the screen, as was the number of seconds left in the current work period.

The slider task was based of a task previously used by \citet{Gill2012}. In each
period of the slider task, five horizontal sliders appeared on the screen (Figure~\ref{fig:taskpictures}a). Each
started at a random setting between 0 and 100, with the slider's value
shown to its right, with a random horizontal offset so that the
sliders were not aligned. The participant's task was to use the mouse to move
each slider to ``50'' before the work period ended. When a participant released
the mouse at the correct setting, the slider turned green to show it had been
completed. To ensure that the task took close to the allotted 30 seconds, at the
beginning of the task only the top slider was enabled, and the other four were
grayed out. Additional sliders were enabled at five-second intervals, such that
all five sliders were available after 20 seconds.

Before beginning the experiment, participants chose one of four videos available
on Youtube: an episode of ``Planet Earth'', and episode of ``The Great British
Bakeoff'', and episode of ``Mythbusters'', or an Ellen Degeneres comedy special.
The video was embedded in the experiment with all user controls (such as
skipping ahead) disabled (Figure~\ref{fig:taskpictures}b). When given access to the video, participants had to keep the
browser window open and hold down the space bar for the video to play. This allowed us to ensure that
participants maintained engagement with the content.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/machinemappictures.png}
\caption{The machine display seen by participants to allow them to keep track of
  the value of each of the six machines and the next time it would be ready to
  make a choice or produce an outcome. Gray arrows have been added to depict the
  counterclockwise movement of machines around the display after each work
  period.
  \emph{a}: the display seen by participants in Experiment 1 and the immediate condition of Experiment 2.
  \emph{b}: the display seen by participants in the delayed condition of
  Experiment 2.}
\label{fig:machinemappictures}
\end{figure*}

Participants completed a total of 70 work
periods. For the first 10 work periods, participants simply clicked a button to begin the
slider task. After these initial periods, participants gained access to six
machines that could potentially complete the slider task for the participant,
allowing the participant to watch their chosen video instead. However, the
machines did not always function, and participants had to make a decision about
how to set the machine before each use.
% , each of which had a ``current spinner'' (exploit) and a ``new
% spinner'' (explore; Figure~\ref{fig:taskpictures}\emph{c}).
% Before each work period, participants
% were shown a machine and made a choice about which spinner to run. The spinner
% then spun, as described below, and if it landed on a ``gold'' wedge then the machine performed the work task for the participant,
%  allowing the participant to watch their chosen video instead.

\emph{Decision-making task.}
Following the initial 10 periods, participants were shown a machine before each work period and had to select
between two circular spinners with arrows at the top, as shown in Figure~\ref{fig:taskpictures}c: the ``current spinner''
(exploit) and the ``new spinner'' (explore).
The current spinner was split into
five black and five gold wedges. If a participant chose the
current spinner, it spun (immediately or after four periods, depending
on the condition) and, if the arrow landed on gold, the machine
worked and the participant could watch the video. Initially, the
current spinner's gold wedges were randomly set for each machine to cover between
1/3 and 2/3 of the spinner.

The new spinner initially showed a question mark. If a participant chose the
new spinner, then a new spinner was
created and appeared on the machine. In the high-explore condition, the new spinner's gold wedges could cover
anywhere from 25\% of the spinner to 100\% of the spinner. In the low-explore condition, the new spinner's gold wedges could cover
anywhere from 0\% of the spinner to 75\% of the spinner. The new spinner then
spun and, if the arrow landed on gold, the machine worked.

Critically, if the new spinner had a greater gold area than the current
spinner, the new spinner was ``saved'' and the current spinner was
updated to the new spinner.  This created an explore--exploit tradeoff in which 
choosing a new spinner carried immediate risk, but could carry long-term
benefits by improving the current spinner from its initial value.
% Importantly, the value of choosing a new spinner was higher in the high-spinner
% condition, because the range of outcomes was shifted towards higher values.

The machine ran immediately after the choice was
made, and affected the next work period, as shown in
Figure~\ref{fig:machinetimeline}a and Figure~\ref{fig:machinemappictures}a. It then ``cooled off'' for the
following five periods, as choices were made with the other five machines.

@@

\emph{Training, incentives, and post-experiment questions.}
Before beginning the full experiment, participants completed two
practice phases. First, they performed several trials of practice using the
machines, with the actual work periods removed. Then, they performed three work
periods practicing the slider task.
During the machine choices, participants had access to an ``info'' button at the
bottom of the screen that provided reminders about the dynamics of the task.

@@

These results show that participants' behavior in this task is
similar to behavior in other exploratory choice tasks. Participants explored more
when the current exploitative option is poor, as well as when the possible
outcomes of exploration are good. While this experiment did not address whether the
consumption rewards produced a bias towards the present (a topic returned to in the
discussion), it did showed that participants enjoyed the video more than the slider
task, and that they made choices to obtain more video time. In the following
experiment, we introduce a delayed-outcome version of the task and predict that, if the
consumption rewards induce a present bias, this version will increase exploration.

@@

Forty participants completed the experiment, which was conducted over
AMT using the psiTurk framework.
Participants were paid \$5.00 for their participation, with a performance-based
bonus of up to \$3.00. All participants received the full \$3.00 bonus.
Participants were pseudo-randomly counterbalanced across two conditions.

@@

Experiment 2 followed the design of Experiment 1, with the following changes. In
both conditions, the possible winning proportion of a new spinner ranged from 0\% to 100\% of
the spinner. In the immediate condition, the machine ran immediately after the
choice was made, as in Experiment 1. In the delayed condition, each machine was
presented to the participant four work periods before it was schedules to run,
and the participant made a choice at that time. The machine then had to
``process'' for four work periods, thus delaying
the outcome by over 2 minutes Figure~\ref{fig:machinemappictures}b. The participant then returned to the
machine to observe its outcome and either perform the slider task or
watch the video. The machine then cooled of for a single period before being
ready for another choice. The timeline for the delayed condition is shown in
Figure~\ref{fig:machinetimeline}b.

@@

\begin{small}
  \noindent
\textbf{Acknowledgments} 
\hl{FUNDING?}
\end{small}
